Clicks and Cliques

### 2018-12-28 papersoul.Rnw
update bib and names

\author{Natalia da Silva \and
        Ignacio Alvarez
}

@@

\institute{Natalia da Silva \at
              Department of Statistics, Iowa State University 
              \email{natalia@iesta.edu.uy}
           \and
           Ignacio Alvarez \at
          Department of Statistics, Iowa State University 
           \email{nachalca@iesta.edu.uy}
           %S. Author \at
            %  second address
}

@@

\keywords{community attachment \and exploratory data analysis \and random forests \and product plots}
\end{abstract}

@@

# stat vizualization
library(productplots) 
library(ggplot2)
library(GGally)
library(maps)

# stat methods
library(randomForest)
library(survey)
library(fpc)
library(reldist) 
library(ggdendro)

@@

Sys.setenv(TEXINPUTS = getwd(),
           BIBINPUTS = getwd(),
           BSTINPUTS = getwd())
@

@@

@@

\subsection{Data description}
The main data set used in this paper was provided by the Knight Foundation, consisting of collected data from 47821 people over three years (2008, 2009 and 2010) in 26 communities across the United States. The data is available at http://streaming.stat.iastate.edu/dataexpo/2013/ as part of the 2013 Data Expo competition of the American Statistical Association. Additional details on the 2013 Data Expo can be found in \cite{soul2}.

@@

As mentioned earlier, information from ACS corresponds to tables of comunity totals (e.g. Total of 18 years old women in each community). These variables were already expanded by Census Bureau, so there is no need to consider sampling weights again. For instance, proportions of age group by gender shown in Figrue \ref{age} are computed directly from ACS data without explicit use of sampling weights.

When using data from Knight Foundation surveys, the sampling weights need to be taken into
account. There are two variables representing sampling weights that might be used considering if data from many communities are combined (PROJWT) or computations are done for each community separatedly (WEIGHT) \citep{full2010}. For instance, Figure \ref{pp1} includes data from all communities toghether so PROJWT is used as sampling weight, meanwhile Figure \ref{comu} shows proportions of attachment computed separatedly within each community  and therefore sampling weights correspond to WEIGHT variable. Additionally, when all three year survey data are used jointly, sampling weights are divided by three so the total aggregate of each variable still can be consider an estimate of the comunity total.

@@

@
\caption{Parallel coordinate plot for community cluster means. \label{cluster.cen} }
\end{figure}

@@

A parallel coordinate plot \citep{inselberg1985plane} is used to compare the cluster means for the variables used in the analysis. Figure \ref{cluster.cen} shows the variables used in the horizontal axis and the variable mean of each cluster on the vertical axis. Group 1 are communities with older people, low education level and biggest home owner rate. Group 2 is formed by communities with large proportion of African-Americans and unequal income distribution. Bradenton and Palm Beach (entire group 3) are communities with older people, not working and not African-American. Group 4 is formed by four communities: Charlotte, Grand Forks, Long Beach, State College and Tallahassee, all relatively younger people, with high education levels and non equitable income distribution. Group 3 and 4 present opposite values in most analyzed variables. For example Age.Fem, Unemp and No.workers are positive and have the biggest values for group 3 among all the groups while for group 4 those variables have the smallest values among all groups. Groups 1, 2 and 3 are similar in Yr.1821, Bachelor , Gini and New.entry and are different in this variabels to group 4.

@@

@
\caption{Cluster mean points. Mean values for all the variables used to create the clusters in every year. \label{para}}
\end{figure}

@@

@
\caption{Attachment vs economic dimension. Top panel shows the correlation between attachment and economy index against time. Bottom panel shows the GDP growth rate against years. \label{eco}}
\end{figure}
\newpage

@@

The Blau index is defined as $D = 1- \sum p_i^2$ where $p_i$ is the proportion of the $i$-th group in the community.  This measure coincides with the Gini impurity measure used in statistics for decision tree related models\footnote{At the same time, Gini impurity measure is different than the Gini index used for measure income distribution in Section \ref{descri}} (as random forest).
\begin{figure}

@@

data.dendro %>% segment() %>% 
  ggplot() + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend),extent = "device") +
  geom_text(data = ggdendro::label(data.dendro), aes(label = label, x = x, y = -1.5)) + 
  coord_flip() + scale_y_reverse() + 
  theme(legend.position = 'none',axis.title.x = element_blank(),axis.title.y = element_blank()) + ylim(c(10,-3))
@
\caption{Dendrogam of communities based on importance measures. Communities close to each other have similar structure of key drivers of attachment. \label{dendro}}
\end{figure}

>>>>

\author{Natalia da Silva \and
        Ignacio Alvarez-Castro
}

@@

\institute{Natalia da Silva \at
              Department of Statistics, Iowa State University 
              \email{natalia@iesta.edu.uy}
           \and
           Ignacio Alvarez-Castro \at
          Department of Statistics, Iowa State University 
           \email{nachalca@iesta.edu.uy}
           %S. Author \at
            %  second address
}

@@

\keywords{community attachment \and exploratory data analysis \and random forests \and statistical visualization \and mosaic plot \and product plot}
\end{abstract}

@@

# stat vizualization
library(productplots) 
library(ggdendro)
library(GGally)

# stat methods
library(randomForest)
library(survey)
library(reldist)

@@

@

@@

This paper was part of 2013 Data Expo competition of the American Statistical Association. An overview of all papers and the data exposition can be
found in \cite{soul2}. A reproducible version of this paper is placed in the following  In GitHub repository: https://github.com/COSTDataExpo2013/DaSilvaAlvarez.

@@

The results in this paper are obtained using R language \citep{Rlang}, and the following {\verb R } packages: {\verb tidyverse } \citep{pkgtidyverse}, {\verb car } \citep{pkgcar}, {\verb GGally } \citep{pkgGGally}, {\verb productplots } \citep{pkgproductplots}, {\verb ggdendro } \citep{pkgggdendro}, {\verb randomForest } \citep{pkgrandomForest}, {\verb survey } \citep{pkgsurvey}, {\verb reldist } \citep{pkgreldist}, {\verb xtable } \citep{pkgxtable} and {\verb knitr } \cite{pkgknitr}.

\subsection{Data description}
The main data set used in this paper was provided by the Knight Foundation, consisting of collected data from 47821 people over three years (2008, 2009 and 2010) in 26 communities across the United States. The data is available at http://streaming.stat.iastate.edu/dataexpo/2013/ as part of the 2013 Data Expo competition of the American Statistical Association.

@@

As mentioned earlier, information from ACS corresponds to tables of community totals (e.g. Total of 18 years old women in each community). These variables were already expanded by Census Bureau, so there is no need to consider sampling weights again. For instance, proportions of age group by gender shown in Figure \ref{age} are computed directly from ACS data without explicit use of sampling weights.

When using data from Knight Foundation surveys, the sampling weights need to be taken into
account. There are two variables representing sampling weights that might be used considering if data from many communities are combined (PROJWT) or computations are done for each community separately (WEIGHT) \citep{full2010}. For instance, Figure \ref{pp1} includes data from all communities together so PROJWT is used as sampling weight, meanwhile Figure \ref{comu} shows proportions of attachment computed separately within each community  and therefore sampling weights correspond to WEIGHT variable. Additionally, when all three year survey data are used jointly, sampling weights are divided by three so the total aggregate of each variable still can be consider an estimate of the community total.

@@

@
\caption{Parallel coordinate plot for community cluster means, color indicates each cluster. Group 3 and 4 present opposite values in most analyzed variables. \label{cluster.cen} }
\end{figure}

@@

A parallel coordinate plot \citep{inselberg1985plane} is used to compare the cluster means for the variables used in the analysis. Figure \ref{cluster.cen} shows the variables used in the horizontal axis and the variable mean of each cluster on the vertical axis. Group 1 are communities with older people, low education level and biggest home owner rate. Group 2 is formed by communities with large proportion of African-Americans and unequal income distribution. Bradenton and Palm Beach (entire group 3) are communities with older people, not working and not African-American. Group 4 is formed by four communities: Charlotte, Grand Forks, Long Beach, State College and Tallahassee, all relatively younger people, with high education levels and non equitable income distribution.

@@

@
\caption{Cluster mean points. Mean values for all the variables used to create the clusters in every year. Color and shape represents each cluster. Clusters 3 and 4 differs only in clique variables (family and friends). \label{para}}
\end{figure}

@@

@
\caption{Attachment vs economic dimension. Bottom panel shows the correlation between attachment and economy index against time. Top panel shows the GDP growth rate against years. \label{eco}}
\end{figure}
\newpage

@@

The Blau index is defined as $D = 1- \sum p_i^2$ where $p_i$ is the proportion of the $i$-th group in the community.  This measure coincides with the Gini impurity measure used in statistics for decision tree related models (as random forest). At the same time, Gini impurity measure is different than the Gini index used for measure income distribution in Section \ref{descri}.
\begin{figure}

@@

data.dendro %>% segment() %>% 
  ggplot() + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend),extent = "device") +
  geom_text(data = ggdendro::label(data.dendro), aes(label = label, x = x, y = -1.5)) + 
  coord_flip() + scale_y_reverse() + 
  theme(legend.position = 'none',axis.title.x = element_blank(),axis.title.y = element_blank()) + ylim(c(10,-3))
@
\caption{Dendrogram of communities based on importance measures. Communities close to each other have similar structure of key drivers of attachment. \label{dendro}}
\end{figure}
